---
title: "Worldbank Agriculture Analysis"
author: "Robin Billinger"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

## Introduction
This analysis explores relationships between indicators across countries such as percentage of agricultural land, CO2 emissions per capita, and size of surface area using World Bank data. It is divided into two main parts:  

1. Is there a relationship between the percentage of agricultural land and CO2 emissions per capita across countries?
2. Does the size of the surface area of the country play a role?.
<br>
<br>

```{r setup, message = FALSE, echo = FALSE}

# Load the processed "data_merged" dataset
data_agriculture <- readRDS(normalizePath("Data/Processed/data_merged.RDS"))
# Filter relevant columns
data_agriculture <- data_agriculture %>%
  select(`Country Name`, `Country Code`, Year, AG.LND.AGRI.ZS, AG.SRF.TOTL.K2, EN.GHG.CO2.MT.CE.AR5)
# Convert column `Year` to data type Integer
data_agriculture$Year <- as.integer(data_agriculture$Year)
```

### 1. Percentage of agricultural land and CO2 emissions per capita
We analyze whether the percentage of agricultural land relates to the CO2 emissions per capita. To do so, we first explore the interested data by looking at the two variables separately.
<br>
<br>
```{r scatter-debt-education, echo = FALSE, , warning=FALSE, message=FALSE}
# Scatter plot: Central government debt vs labor force with basic education
ggplot(data_education, aes(x = GC.DOD.TOTL.GD.ZS, y = SL.TLF.BASC.ZS, color = income_level)) +
  geom_point(alpha = 0.7, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(
    title = "Central Government Debt vs Labor Force with Basic Education",
    x = "Central Government Debt (% of GDP)",
    y = "Labor Force with Basic Education (%)",
    color = "Income Level"
  ) +
  theme_minimal()
```
