---
title: "Worldbank HIV Analysis"
author: "Leonie Mertes"
date: "`r Sys.Date()`"
output:
  html_document: default
---

## Introduction
This analysis explores relationships of HIV prevalence in the population aged 15-49 to the total alcohol consumption per capita as well as its coherence to percentage of labor force with basic education. Therefore we have two main parts:
1. How does the HIV prevalence in the population aged 15-49 relate to the total alcohol consumption per capita?
2. Do countries with higher percentage of labor force with basic education have lower HIV prevalence rates in the 15-49 population?
<br>
<br>
```{r setup, message = FALSE, echo = FALSE, warning = FALSE}
# Load necessary libraries
library(dplyr)
library(ggplot2)

# Load the dataset
data_merged <- readRDS("C:/Users/Leonie/Documents/Studium/3. Semester/Statsoft/worldbank_analysis/Data/Processed/data_merged.RDS")
data_grouped <- readRDS("C:/Users/Leonie/Documents/Studium/3. Semester/Statsoft/worldbank_analysis/Data/Processed/data_grouped.RDS")
data_processed <- left_join(data_merged, data_grouped, by = c("Country Name", "Country Code"))
data_hiv <- data_processed %>% 
  select(`Country Name`, `Country Code`, Year, SH.DYN.AIDS.ZS, SH.ALC.PCAP.LI, SL.TLF.BASC.ZS, alc_consumption, edu_basic)
```
<br>
Check how many NA's in HIV, alcohol and education variable each Year (excluding years existing only of NA's)
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  group_by(Year) %>%
  reframe(
    num_na_hiv = sum(is.na(SH.DYN.AIDS.ZS)),
    num_na_alc = sum(is.na(SH.ALC.PCAP.LI)),
    num_na_edu = sum(is.na(SL.TLF.BASC.ZS))
  ) %>%
  filter(num_na_hiv < 25 & num_na_alc < 25 & num_na_edu < 25)
```
<br>  
### 1. HIV Prevalence and alcohol consumption
<br>
To take a look at the grouping of our data by alcohol consumption, we will display a plot of the most recent data in Year 2020.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  ggplot(aes(`Country Name`, SH.ALC.PCAP.LI, colour = alc_consumption)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    y = "Alkoholkonsum in Litern pro Kopf",
    x = "Land",
    title = "Gruppierung der Daten im Jahr 2020",
    colour = "verhältnismäßiger Alkoholkonsum"
  ) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 4
  )) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Maybe its possible to group data for each year?
<br>
Now we take a look at the Comparison of alcohol consumption and HIV prevalence in 2020. Therefore we group by the relative alcohol consumption.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  group_by(alc_consumption) %>%
  ggplot(aes(SH.ALC.PCAP.LI, SH.DYN.AIDS.ZS, colour = alc_consumption)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Alkoholkonsum in Litern pro Kopf",
    y = "HIV Prävalenz in %",
    title = "Alkoholkonsum vs. HIV Prävalenz im Jahr 2020",
    colour = "relativer Alkoholkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
In the following we view the same plot, but without the extreme value. Therefore the country Tanzania is removed.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  filter(`Country Name` != "Tanzania") %>%
  group_by(alc_consumption) %>%
  ggplot(aes(SH.ALC.PCAP.LI, SH.DYN.AIDS.ZS, colour = alc_consumption)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Alkoholkonsum in Litern pro Kopf",
    y = "HIV Prävalenz in %",
    title = "Alkoholkonsum vs. HIV Prävalenz im Jahr 2020",
    colour = "relativer Alkoholkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now we view only one dot per category of relative alcohol consumption
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  group_by(alc_consumption) %>%
  summarise(
    alc_consumption = alc_consumption,
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SH.ALC.PCAP.LI = mean(SH.ALC.PCAP.LI, na.rm = TRUE)
  ) %>%
  ggplot(aes(SH.ALC.PCAP.LI, SH.DYN.AIDS.ZS, colour = alc_consumption)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    title = "Alkoholkonsum vs. HIV Prävalenz im Jahr 2020",
    x = "Alkoholkonsum in Litern pro Kopf",
    y = "HIV Prävalenz",
    colour = "relativer Alkoholkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now again without the extreme value of Tanzania
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  filter(`Country Name` != "Tanzania") %>%
  group_by(alc_consumption) %>%
  summarise(
    alc_consumption = alc_consumption,
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SH.ALC.PCAP.LI = mean(SH.ALC.PCAP.LI, na.rm = TRUE)
  ) %>%
  ggplot(aes(SH.ALC.PCAP.LI, SH.DYN.AIDS.ZS, colour = alc_consumption)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    title = "Alkoholkonsum vs. HIV Prävalenz im Jahr 2020",
    x = "Alkoholkonsum in Litern pro Kopf",
    y = "HIV Prävalenz",
    colour = "relativer Alkoholkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
To get a better overview we take a look at all years.
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  group_by(Year, alc_consumption) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SH.ALC.PCAP.LI = mean(SH.ALC.PCAP.LI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(Year, SH.DYN.AIDS.ZS, colour = alc_consumption)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(aes(group = alc_consumption)) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 6
  )) +
  labs(
    title = "HIV Prävalenz über die Jahre",
    x = "Jahr",
    y = "HIV Prävalenz",
    colour = "relativer Alkoholkonsum"
  ) +
  scale_colour_brewer(palette = "Set2")
```
<br>
(bringt nichts!:
Now we want to take a more precise look at the connection of HIV Prevalence and alcohol consumption. Therefore we facet over year.)
<br>
<br>
### 2. HIV prevalence and basic education
<br>
To take a look at the grouping of our data by basic education of the working force, we will display a plot of the most recent data in Year 2021.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2021) %>%
  ggplot(aes(`Country Name`, SL.TLF.BASC.ZS, colour = edu_basic)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    y = "Anteil an Arbeitenden mit grundlegender Bildung",
    x = "Land",
    title = "Gruppierung der Daten im Jahr 2021",
    colour = "verhältnismägies Bildungslevel"
  ) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 4
  )) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now we take a look at the Comparison of percentage of the labor force with basic education and HIV prevalence in 2021. Therefore we group by the relative education level.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2021) %>%
  group_by(edu_basic) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = edu_basic)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil an Arbeitenden mit grundlegender Bildung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Jahr 2021",
    colour = "relativer Bildungsstand"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
This plot is not very meaningful, therfore we try to only plot one dot per category.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2021) %>%
  group_by(edu_basic) %>%
  summarise(
    edu_basic = edu_basic,
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SL.TLF.BASC.ZS = mean(SL.TLF.BASC.ZS, na.rm = TRUE)
  ) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = edu_basic)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil an Arbeitenden mit grundlegender Bildung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Jahr 2021",
    colour = "relativer Bildungsstand"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2") 
```
<br>
Now we see a upwards tendency (even without the extreme value).
We notice as well that there is no value for the category "Medium" relative education. Therefore we take a look at 2020.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  group_by(edu_basic) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = edu_basic)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil an Arbeitenden mit grundlegender Bildung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Jahr 2020",
    colour = "relativer Bildungsstand"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Is there really an downwards tendency?
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  group_by(edu_basic) %>%
  summarise(
    edu_basic = edu_basic,
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SL.TLF.BASC.ZS = mean(SL.TLF.BASC.ZS, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = edu_basic)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil an Arbeitenden mit grundlegender Bildung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Jahr 2020",
    colour = "relativer Bildungsstand"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Unfortunately in mean the graph is still upwards. Maybe plot over years to get an overview.
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  group_by(Year, edu_basic) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SL.TLF.BASC.ZS = mean(SL.TLF.BASC.ZS, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(Year, SH.DYN.AIDS.ZS, colour = edu_basic)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(aes(group = edu_basic)) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 6
  )) +
  labs(
    title = "HIV Prävalenz über die Jahre",
    x = "Jahr",
    y = "HIV Prävalenz",
    colour = "relativer Bildungsstand"
  )  +
  scale_colour_brewer(palette = "Set2")
```
<br>
The 60 % with higher education level also show a higher HIV Prevalence. Their Prevalence is falling over the years, while countries with lower education level have an small tendency to increasing HIV prevalence.
