---
title: "Worldbank HIV Analysis"
author: "Leonie Mertes"
date: "`r Sys.Date()`"
output:
  html_document: default
---

## Introduction
This analysis explores relationships of HIV prevalence in the population aged 15-49 to the total alcohol consumption per capita as well as its coherence to percentage of labor force with basic education. Therefore we have two main parts:
1. How does the HIV prevalence in the population aged 15-49 relate to the total alcohol consumption per capita?
2. Do countries with higher percentage of labor force with basic education have lower HIV prevalence rates in the 15-49 population?
<br>
<br>
```{r setup, message = FALSE, echo = FALSE, warning = FALSE}
# Load necessary libraries
library(here)
library(dplyr)
library(ggplot2)
library(tidyr)

# Load the dataset
data_merged <- readRDS(here("Data", "Processed", "data_merged.RDS"))
data_grouped <- readRDS(here("Data", "Processed", "data_grouped.RDS"))
data_processed <- left_join(data_merged, data_grouped, by = c("Country Name", "Country Code"))
data_hiv <- data_processed %>% 
  select(`Country Name`, `Country Code`, Year, SH.DYN.AIDS.ZS, SH.ALC.PCAP.LI, SL.TLF.BASC.ZS, cat_alc, cat_education)
```
<br>
Check how many NA's in HIV, alcohol and education variable each Year (excluding years existing only of NA's)
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  group_by(Year) %>%
  reframe(
    num_na_hiv = sum(is.na(SH.DYN.AIDS.ZS)),
    num_na_alc = sum(is.na(SH.ALC.PCAP.LI)),
    num_na_edu = sum(is.na(SL.TLF.BASC.ZS))
  ) %>%
  filter(num_na_hiv < 25 & num_na_alc < 25 & num_na_edu < 25)
```
<br>  
### 1. HIV Prevalence and alcohol consumption
<br>
To take a look at the grouping of our data by alcohol consumption, we will display a plot of the most recent data in Year 2020.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  ggplot(aes(`Country Name`, SH.ALC.PCAP.LI, colour = cat_alc)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    y = "Alkoholkonsum in Litern pro Kopf",
    x = "Land",
    title = "Gruppierung der Daten im Jahr 2020",
    colour = "verhältnismäßiger Alkoholkonsum"
  ) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 4
  )) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Maybe its possible to group data for each year?
<br>
Now we take a look at the Comparison of alcohol consumption and HIV prevalence in 2020. Therefore we group by the relative alcohol consumption.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  group_by(cat_alc) %>%
  ggplot(aes(SH.ALC.PCAP.LI, SH.DYN.AIDS.ZS, colour = cat_alc)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Alkoholkonsum in Litern pro Kopf",
    y = "HIV Prävalenz in %",
    title = "Alkoholkonsum vs. HIV Prävalenz im Jahr 2020",
    colour = "relativer Alkoholkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
In the following we view the same plot, but without the extreme value. Therefore the country Tanzania is removed.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  filter(`Country Name` != "Tanzania") %>%
  group_by(cat_alc) %>%
  ggplot(aes(SH.ALC.PCAP.LI, SH.DYN.AIDS.ZS, colour = cat_alc)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Alkoholkonsum in Litern pro Kopf",
    y = "HIV Prävalenz in %",
    title = "Alkoholkonsum vs. HIV Prävalenz im Jahr 2020",
    colour = "relativer Alkoholkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now we view only one dot per category of relative alcohol consumption.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  group_by(cat_alc) %>%
  summarise(
    cat_alc = cat_alc,
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SH.ALC.PCAP.LI = mean(SH.ALC.PCAP.LI, na.rm = TRUE)
  ) %>%
  ggplot(aes(SH.ALC.PCAP.LI, SH.DYN.AIDS.ZS, colour = cat_alc)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    title = "Alkoholkonsum vs. HIV Prävalenz im Jahr 2020",
    x = "Alkoholkonsum in Litern pro Kopf",
    y = "HIV Prävalenz",
    colour = "relativer Alkoholkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now again without the extreme value of Tanzania
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  filter(`Country Name` != "Tanzania") %>%
  group_by(cat_alc) %>%
  summarise(
    cat_alc = cat_alc,
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SH.ALC.PCAP.LI = mean(SH.ALC.PCAP.LI, na.rm = TRUE)
  ) %>%
  ggplot(aes(SH.ALC.PCAP.LI, SH.DYN.AIDS.ZS, colour = cat_alc)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    title = "Alkoholkonsum vs. HIV Prävalenz im Jahr 2020",
    x = "Alkoholkonsum in Litern pro Kopf",
    y = "HIV Prävalenz",
    colour = "relativer Alkoholkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
To get a better overview we take a look at all years.
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  group_by(Year, cat_alc) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SH.ALC.PCAP.LI = mean(SH.ALC.PCAP.LI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(Year, SH.DYN.AIDS.ZS, colour = cat_alc)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(aes(group = cat_alc)) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 6
  )) +
  labs(
    title = "HIV Prävalenz über die Jahre",
    x = "Jahr",
    y = "HIV Prävalenz",
    colour = "relativer Alkoholkonsum"
  ) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now we take a look over the average of all years, where we categorize by relative alcohol consumption and compare to HIV Prevalence.
<br> 
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  group_by(cat_alc) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = cat_alc, y = SH.DYN.AIDS.ZS)) +
  geom_col(fill = "steelblue") +
  labs(
    x = "relativer Alkoholkonsum",
    y = "HIV Prävalenz in %",
    title = "Alkoholkonsum vs. HIV Prävalenz im Schnitt über die Jahre"
  )
```
<br>
to take a look at the categorization, we plot the same data in slightly different way:
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  group_by(cat_alc) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SH.ALC.PCAP.LI = mean(SH.ALC.PCAP.LI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(SH.ALC.PCAP.LI, SH.DYN.AIDS.ZS, colour = cat_alc)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Alkoholkonsum in Litern pro Kopf",
    y = "HIV Prävalenz in %",
    title = "Alkoholkonsum vs. HIV Prävalenz im Schnitt über die Jahre",
    colour = "relativer Alkoholkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
### 2. HIV prevalence and basic education
<br>
To take a look at the grouping of our data by basic education of the working force, we will display a plot of the most recent data in Year 2021.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2021) %>%
  ggplot(aes(`Country Name`, SL.TLF.BASC.ZS, colour = cat_education)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    y = "Anteil an Arbeitenden mit grundlegender Bildung",
    x = "Land",
    title = "Gruppierung der Daten im Jahr 2021",
    colour = "verhältnismägies Bildungslevel"
  ) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 4
  )) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now we take a look at the Comparison of percentage of the labor force with basic education and HIV prevalence in 2021. Therefore we group by the relative education level.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2021) %>%
  group_by(cat_education) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = cat_education)) +
  geom_point(size = 3, alpha = 0.7) +
  coord_cartesian(ylim = c(0, 1.75)) +
  labs(
    x = "Anteil an Arbeitenden mit grundlegender Bildung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Jahr 2021",
    colour = "relativer Bildungsstand"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
This plot is not very meaningful, therfore we try to only plot one dot per category.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2021) %>%
  group_by(cat_education) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SL.TLF.BASC.ZS = mean(SL.TLF.BASC.ZS, na.rm = TRUE)
  ) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = cat_education)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil an Arbeitenden mit grundlegender Bildung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Jahr 2021",
    colour = "relativer Bildungsstand"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2") 
```
<br>
Now we see a upwards tendency (even without the extreme value).
We notice as well that there is no value for the category "Medium" relative education. Therefore we take a look at 2020.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  group_by(cat_education) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = cat_education)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil an Arbeitenden mit grundlegender Bildung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Jahr 2020",
    colour = "relativer Bildungsstand"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Is there really an downwards tendency?
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>%
  group_by(cat_education) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SL.TLF.BASC.ZS = mean(SL.TLF.BASC.ZS, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = cat_education)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil an Arbeitenden mit grundlegender Bildung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Jahr 2020",
    colour = "relativer Bildungsstand"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Unfortunately in mean the graph is still upwards. Maybe plot over years to get an overview.
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  group_by(Year, cat_education) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SL.TLF.BASC.ZS = mean(SL.TLF.BASC.ZS, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(Year, SH.DYN.AIDS.ZS, colour = cat_education)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(aes(group = cat_education)) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 6
  )) +
  labs(
    title = "HIV Prävalenz über die Jahre",
    x = "Jahr",
    y = "HIV Prävalenz",
    colour = "relativer Bildungsstand"
  )  +
  scale_colour_brewer(palette = "Set2")
```
<br>
The 60 % with higher education level also show a higher HIV Prevalence. Their Prevalence is falling over the years, while countries with lower education level have an small tendency to increasing HIV prevalence. The labor force with relativly medium
education have a higher HIV Prevalence than the others.
<br>
Now we take a look at all years, facetted by cat_education, all countries each a line
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  filter(Year == 2020) %>% 
  group_by(cat_education) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = cat_education)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(aes(group = cat_education)) +
  labs(
    title = "HIV Prävalenz im Jahr 2020",
    x = "Anteil grundlegend gebildeter Arbeitskräfte",
    y = "HIV Prävalenz",
    colour = "relativer Bildungsstand"
  )  +
  scale_colour_brewer(palette = "Set2")
```
<br>
we want to create a plot of the averages over alll years grouped by education level. For interest first with x-axis education and second as bar plot (x axis the education level)
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  group_by(cat_education) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SL.TLF.BASC.ZS = mean(SL.TLF.BASC.ZS, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = cat_education)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil an Arbeitenden mit grundlegender Bildung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Schnitt über die Jahre",
    colour = "relativer Bildungsstand"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_hiv %>%
  group_by(cat_education) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = cat_education, y = SH.DYN.AIDS.ZS)) +
  geom_col(fill = "steelblue") +
  labs(
    x = "relatives Bildungslevel der arbeitsfähigen Bevölkerung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Schnitt über die Jahre"
  )
```
<br>