---
title: "Worldbank Analysis"
subtitle: "subtitle"
author: "Robin Billinger, Leonie Mertes, Iman Saffari, Maximilian Frei"
institute: "institute"
date: "2025-01-21"
format: 
  revealjs:
    self-contained: true
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
#source("env_setup.R")

```

## Agenda

::: incremental
1.  Überblick und Terminologie

2.  Datenanalyse

    2.1 Electricity and National Income\
    2.2 Education\
    2.3 HIV\
    2.4 Tobacco\
    2.5 Agriculture

3.  Zusammenfassung und Diskussion
:::

## Überblick

**Worldbank Datensatz**

::: incremental
-   **Erhebungsart**: Logitudinal

-   25 Länder über 22 Jahre (2000 - 2021) untersucht

    -- Weitere Details

-   18 verschiedene Indikatoren
:::

## Terminologie

::: incremental
-   **fachbegriff**: kurze Beschreibung

-   **weiterer fachbegriff**: kurze Beschreibung
:::

## 2.1 Electricity and National Income (1/2) {.smaller}

How does access to electricity correlate with adjusted net national income per capita across different countries?

## 2.1 Electricity and National Income (2/2) {.smaller}

Does this also depend on the country size or the inhabitants?

## 2.2 Education (1/2) {.smaller}

Do countries with **higher central government debt (% of GDP)** have a **lower % of labor force with basic education**?

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)

# Load the dataset
data_merged <- readRDS("/Users/maximilian/Desktop/wb_local/Data/Processed/data_merged.RDS")
data_grouped <- readRDS("/Users/maximilian/Desktop/wb_local/Data/Processed/data_grouped.RDS")
data_education <- left_join(data_merged, data_grouped, by = c("Country Name", "Country Code"))

# Faceted scatter plot: Debt vs Education by Income Level
ggplot(data_education, aes(x = GC.DOD.TOTL.GD.ZS, y = SL.TLF.BASC.ZS, color = cat_income_level)) +
  geom_point(alpha = 0.7, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  facet_wrap(~ cat_income_level) +
  labs(
    title = "Central Government Debt vs Labor Force with Basic Education by Income Level",
    x = "Central Government Debt (% of GDP)",
    y = "Labor Force with Basic Education (%)",
    color = "Income Level"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))
```

## 2.2 Education (2/2) {.smaller}

Are countries with higher percentage of labor force with basic education able to maintain **lower** pupil-teacher ratios, and what impact might this have on education quality?

## 2.3 HIV (1/2) {.smaller}

How does the HIV prevalence in the population aged 15-49 relate to the total alcohol consumption per capita?

## 2.3 HIV (2/2) {.smaller}

Do countries with higher percentage of labor force with basic education have lower HIV prevalence rates in the 15-49 population?

## 2.4 Tobacco (1/2) {.smaller}

How does GDP per capita relate to the prevalence of current tobacco use (% of adults)?

## 2.4 Tobacco (2/2) {.smaller}

Further analysis of tobacco.

## 2.5 Agriculture (1/2) {.smaller}

Is there a relationship between the percentage of agricultural land and CO2 emissions per capita across countries?

## 2.5 Agriculture (2/2) {.smaller}

Does the size of the surface area of the country play a role?

## Zusammenfassung und Diskussion {.smaller}
