---
title: "Worldbank Analysis"
subtitle: "subtitle"
author: "Robin Billinger, Maximilian Frei, Leonie Mertes, Iman Saffari"
institute: "institute"
date: "2025-01-20"
format: 
  revealjs:
    self-contained: true
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
#source("env_setup.R")

```

## Agenda

::: incremental
1.  Überblick und Terminologie

2.  Datenanalyse

    2.1 Elektrizität und nationales Einkommen\
    2.2 Bildung\
    2.3 HIV\
    2.4 Tabakkonsum\
    2.5 Landwirtschaft

3.  Zusammenfassung und Diskussion
:::

## Überblick

**Worldbank Datensatz**

::: incremental
-   **Erhebungsart**: Longitudinal

-   25 Länder über 22 Jahre (2000 - 2021) untersucht

    -- Weitere Details

-   18 verschiedene Indikatoren
:::

## Terminologie

::: incremental
-   **fachbegriff**: kurze Beschreibung

-   **weiterer fachbegriff**: kurze Beschreibung
:::

## 2.1 Electricity and National Income (1/2) {.smaller}

How does access to electricity correlate with adjusted net national income per capita across different countries?

## 2.1 Electricity and National Income (2/2) {.smaller}

Does this also depend on the country size or the inhabitants?

## 2.2 Education (1/2) {.smaller}

Do countries with **higher central government debt (% of GDP)** have a **lower % of labor force with basic education**?

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)

# Load the dataset
data_merged <- readRDS("/Users/maximilian/Desktop/wb_local/Data/Processed/data_merged.RDS")
data_grouped <- readRDS("/Users/maximilian/Desktop/wb_local/Data/Processed/data_grouped.RDS")
data_education <- left_join(data_merged, data_grouped, by = c("Country Name", "Country Code"))

# Faceted scatter plot: Debt vs Education by Income Level
ggplot(data_education, aes(x = GC.DOD.TOTL.GD.ZS, y = SL.TLF.BASC.ZS, color = cat_income_level)) +
  geom_point(alpha = 0.7, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  facet_wrap(~ cat_income_level) +
  labs(
    title = "Central Government Debt vs Labor Force with Basic Education by Income Level",
    x = "Central Government Debt (% of GDP)",
    y = "Labor Force with Basic Education (%)",
    color = "Income Level"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))
```

## 2.2 Education (2/2) {.smaller}

Are countries with higher percentage of labor force with basic education able to maintain **lower** pupil-teacher ratios, and what impact might this have on education quality?

## 2.3 HIV (1/2) {.smaller}

How does the **HIV prevalence in the population aged 15-49** relate to the **total alcohol consumption per capita**?

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)

# Load the dataset
data_merged <- readRDS("C:\\Users\\Leonie\\Documents\\Studium\\3. Semester\\StatSoft\\worldbank_analysis\\Data\\Processed\\data_merged.RDS")
data_grouped <- readRDS("C:\\Users\\Leonie\\Documents\\Studium\\3. Semester\\StatSoft\\worldbank_analysis\\Data\\Processed\\data_grouped.RDS")
data_processed <- left_join(data_merged, data_grouped, by = c("Country Name", "Country Code"))
data_hiv <- data_processed %>% 
  select(`Country Name`, `Country Code`, Year, SH.DYN.AIDS.ZS, SH.ALC.PCAP.LI, SL.TLF.BASC.ZS, cat_alc, cat_education)

data_hiv %>%
  group_by(Year, cat_alc) %>%
  summarise(
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SH.ALC.PCAP.LI = mean(SH.ALC.PCAP.LI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(Year, SH.DYN.AIDS.ZS, colour = cat_alc)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(aes(group = cat_alc)) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 6
  )) +
  labs(
    title = "HIV Prävalenz über die Jahre",
    x = "Jahr",
    y = "HIV Prävalenz",
    colour = "relativer Alkoholkonsum"
  ) +
  scale_colour_brewer(palette = "Set2")
```

## 2.3 HIV (2/2) {.smaller}

Do countries with higher percentage of labor force with basic education have lower HIV prevalence rates in the 15-49 population?

<<<<<<< HEAD
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

# Load the dataset
data_merged <- readRDS("C:\\Users\\Leonie\\Documents\\Studium\\3. Semester\\StatSoft\\worldbank_analysis\\Data\\Processed\\data_merged.RDS")
data_grouped <- readRDS("C:\\Users\\Leonie\\Documents\\Studium\\3. Semester\\StatSoft\\worldbank_analysis\\Data\\Processed\\data_grouped.RDS")
data_processed <- left_join(data_merged, data_grouped, by = c("Country Name", "Country Code"))
data_hiv <- data_processed %>% 
  select(`Country Name`, `Country Code`, Year, SH.DYN.AIDS.ZS, SH.ALC.PCAP.LI, SL.TLF.BASC.ZS, cat_alc, cat_education)

data_hiv %>%
  filter(Year == 2020) %>%
  group_by(cat_education) %>%
  summarise(
    cat_education = cat_education,
    SH.DYN.AIDS.ZS = mean(SH.DYN.AIDS.ZS, na.rm = TRUE),
    SL.TLF.BASC.ZS = mean(SL.TLF.BASC.ZS, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(SL.TLF.BASC.ZS, SH.DYN.AIDS.ZS, colour = cat_education)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil an Arbeitenden mit grundlegender Bildung",
    y = "HIV Prävalenz in %",
    title = "Bildungsstand vs. HIV Prävalenz im Jahr 2020",
    colour = "relativer Bildungsstand"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```

## 2.4 Tabakkonsum (1/2) {.smaller}
>>>>>>> b0148bda1cd10436d14782eb05da10e736db5678

In welchem Zusammenhang steht das Bruttoinlandsprodukt pro Kopf mit der Prävalenz des aktuellen Tabakkonsums unter Erwachsenen innerhalb eines Landes?

```{r tobacco_setup}
# Load the dataset
data_tobacco <- readRDS("Data/Processed/data_merged.RDS")
>>>>>>> b0148bda1cd10436d14782eb05da10e736db5678

# Filter relevant columns Country Name, Country Code, Year, NY.GDP.PCAP.PP.KD and SH.PRV.SMOK 
data_tobacco <- data_tobacco %>%
  select(`Country Name`, `Country Code`, Year, NY.GDP.PCAP.PP.KD, SH.PRV.SMOK)

# Convert column `Year` to data type Integer
data_tobacco$Year <- as.integer(data_tobacco$Year)

# As we only have data on each country for the years 2000, 2005, 2010, 2018, 2019, 2020,
# we remove all missing years for each country
data_tobacco <- data_tobacco[data_tobacco$Year %in% c(2000, 2005, 2010, 2018, 2019, 2020), ]

# Calculate mean and standard deviation for each year
data_tobacco_normdensity <- data_tobacco %>%
  group_by(Year) %>%
  summarize(SH.PRV.SMOK_mean = mean(SH.PRV.SMOK),
            SH.PRV.SMOK_sd = sd(SH.PRV.SMOK))

# Generate evenly-spaced values of x and compute density for each year
data_tobacco_normdensity <- data_tobacco_normdensity %>%
  rowwise() %>%
  mutate(
    # Create 100 evenly-spaced points inbetween the lowest and highest observed datapoint 
    norm_x_val = list(seq(0, 100, length.out = 100)),
    # Compute the corresponding PDF values
    norm_y_val = list(dnorm(norm_x_val, mean = SH.PRV.SMOK_mean, sd = SH.PRV.SMOK_sd))) %>%
  unnest(c(norm_x_val, norm_y_val))  # Convert the lists into columns

# Remove rows with values of the normal distribution calculation that are larger than 60% (as 60% > max(SH.PRV.SMOK))
data_tobacco_normdensity <- data_tobacco_normdensity %>%
  filter(norm_x_val <= 60)
```

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)

# Load the dataset
data_merged <- readRDS(
  "C:/Users/Leonie/Documents/Studium/3. Semester/Statsoft/worldbank_analysis/Data/Processed/data_merged.RDS"
)
data_grouped <- readRDS(
  "C:/Users/Leonie/Documents/Studium/3. Semester/Statsoft/worldbank_analysis/Data/Processed/data_grouped.RDS"
)
data_processed <- left_join(data_merged, data_grouped, by = c("Country Name", "Country Code"))
data_tob <- data_processed %>%
  select(`Country Name`,
         `Country Code`,
         Year,
         NY.GDP.PCAP.PP.KD,
         SH.PRV.SMOK,
         cat_tob_usage)

data_tob %>%
  group_by(Year, cat_tob_usage) %>%
  summarise(
    NY.GDP.PCAP.PP.KD = mean(NY.GDP.PCAP.PP.KD, na.rm = TRUE),
    SH.PRV.SMOK = mean(SH.PRV.SMOK, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(Year, NY.GDP.PCAP.PP.KD, colour = cat_tob_usage)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(aes(group = cat_tob_usage)) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 6
  )) +
  labs(
    x = "Jahr",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP",
    colour = "relativer Tabakkonsum"
  ) +
  scale_colour_brewer(palette = "Set2")
```

## 2.4 Tabakkonsum (2/2) {.smaller}

Tiefergehende Analyse des Tabakkonsums
```{r tobacco_plot2_faceted_scatter_and_density}
# Plot the upper row of the data points for the assembled tobacco plotting
plot_tobacco1_data_upper <- ggplot(data_tobacco %>% filter(Year %in% c(2000, 2005, 2010))) +
  geom_point(aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, group = Year), alpha = 0.5) +  
  ggtitle("Zusammenhang zwischen der Prävalenz des aktuellen Tabakkonsums und BIP pro Kopf") +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]", breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(), limits = c(0, 60)) +
  scale_y_continuous(name = "BIP pro Kopf [$]", breaks = c(0, 25000, 50000, 75000, 100000, 125000),
                     minor_breaks = waiver(), limits = c(0, 125000)) +
  theme_bw() +
  theme(plot.margin = margin(0, 0, 0, 0),
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  facet_wrap(~ Year)

# Plot the bottom row of the data points for the assembled tobacco plotting
plot_tobacco1_data_bottom <- ggplot(data_tobacco %>% filter(Year %in% c(2018, 2019, 2020))) +
  geom_point(aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, group = Year), alpha = 0.5) +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]", breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(), limits = c(0, 60)) +
  scale_y_continuous(name = "BIP pro Kopf [$]", breaks = c(0, 25000, 50000, 75000, 100000, 125000),
                     minor_breaks = waiver(), limits = c(0, 125000)) +
  theme_bw() +
  theme(plot.margin = margin(0, 0, 0, 0),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  facet_wrap(~ Year)

# Add overlayed normal distribution curves to the assembled plot in the data displays
# Plot the upper row of the density curves for the assembled tobacco plotting
plot_tobacco1_density_upper <- ggplot() +
  geom_line(data = data_tobacco_normdensity %>% filter(Year %in% c(2000, 2005, 2010)),  # TODO: data
            aes(x = norm_x_val, y = norm_y_val, group = Year),
            color = brewer.pal(4, "Set2")[[2]], linewidth = 1, inherit.aes = FALSE) +
  geom_density(data = data_tobacco %>% filter(Year %in% c(2000, 2005, 2010)),
               aes(x = SH.PRV.SMOK, y = after_stat(density), group = Year),
               color = brewer.pal(4, "Set2")[[1]], linewidth = 1, alpha = 0.5, bw = 5) +
  geom_vline(data = data_tobacco_normdensity %>% filter(Year %in% c(2000, 2005, 2010)),  # TODO: data
             aes(xintercept = SH.PRV.SMOK_mean, group = Year),
             color = "black", linetype = "dashed") +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]", breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(), limits = c(0, 60)) +
  scale_y_continuous(name = "Dichte", breaks = c(0, 0.02, 0.04), minor_breaks = waiver()) +
  theme_bw() +
  theme(strip.text = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(margin = margin(r = 10))) +
  facet_wrap(~ Year)

# Plot the bottom row of the density curves for the assembled tobacco plotting
plot_tobacco1_density_bottom <- ggplot() +
  geom_line(data = data_tobacco_normdensity %>% filter(Year %in% c(2018, 2019, 2020)),  # TODO: data
            aes(x = norm_x_val, y = norm_y_val, group = Year, color = "Normalverteilung"),
            linewidth = 1, inherit.aes = FALSE) +
  geom_density(data = data_tobacco %>% filter(Year %in% c(2018, 2019, 2020)),
               aes(x = SH.PRV.SMOK, y = after_stat(density), group = Year),
               color = brewer.pal(3, "Set2")[[1]], linewidth = 1, bw = 5, show.legend = FALSE) +
  # Add an invisible line by framing dummy data to show "KDE" with line instead of hollow square for shape in legend
  geom_line(data = data.frame(x = c(0,0), y = c(0,0)),
            aes(x = x, y = y, color = "Kerndichteschätzer"), linewidth = 1) +
  geom_vline(data = data_tobacco_normdensity %>% filter(Year %in% c(2018, 2019, 2020)),  # TODO: data
             aes(xintercept = SH.PRV.SMOK_mean, group = Year),
             color = "black", linetype = "dashed") +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]", breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(), limits = c(0, 60)) +
  scale_y_continuous(name = "Dichte", breaks = c(0, 0.02, 0.04), minor_breaks = waiver()) +
  scale_color_brewer(name = "", palette = "Set2") +
  theme_bw() +
  theme(strip.text = element_blank(),
        axis.title.y = element_text(margin = margin(r = 10)),
        legend.position = "bottom",
        legend.text = element_text(size = 10)) +
  facet_wrap(~ Year)

# Assemble the plot
(plot_tobacco1_data_upper /
    plot_tobacco1_density_upper /
    plot_tobacco1_data_bottom /
    plot_tobacco1_density_bottom) +
  plot_layout(heights = c(2, 0.5, 2, 0.5))
```

## 2.5 Landwirtschaft (1/4) {.smaller}

Gibt es einen Zusammenhang zwischen dem Anteil der landwirtschaftlichen Nutzfläche und den CO2 Emissionen pro Kopf eines Landes?

```{r agriculture_setup_and_plot1_scatter}
# Load the processed "data_merged" dataset
data_agriculture <- readRDS("Data/Processed/data_merged.RDS")

# Filter relevant columns
data_agriculture <- data_agriculture %>%
  select(`Country Name`, `Country Code`, Year, AG.LND.AGRI.ZS, AG.SRF.TOTL.K2, EN.GHG.CO2.MT.CE.AR5)

# Convert column `Year` to data type Integer
data_agriculture$Year <- as.integer(data_agriculture$Year)

# Save the ascending order of average agricultural land [%] by country
country_order_by_agrLand <- data_agriculture %>%
  group_by(`Country Name`) %>%
  summarise(avg_AG.LND.AGRI.ZS = mean(AG.LND.AGRI.ZS)) %>%
  arrange(avg_AG.LND.AGRI.ZS) %>%
  pull(`Country Name`)

# Initial comparison via scatterplot of agricultural land [%] and CO2 emissions per capita [mt]
ggplot(data_agriculture, aes(x = AG.LND.AGRI.ZS, y = log10(EN.GHG.CO2.MT.CE.AR5 + 1))) +
  geom_point(alpha = .3, size = 2.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title = "Zusammenhang zwischen CO2 Emissionen pro Kopf\nund landwirtschaftlicher Nutzfläche",
       x = "Landwirtschaftliche Nutzfläche [%]", y = "Logarithmierte CO2 Emissionen pro Kopf [Mt]") +
  # scale_x_log10(name = "CO2 Emissionen pro Kopf",
  #               breaks = log10(c(0, 10, 100, 1000, 10000, 100000) + 1),
  #               labels = c("0", "10", "100", "1.000", "10.000", "100.000")) +
  # scale_x_log10(breaks = trans_breaks("log10", function(x) 10 ^ x),
  #               labels = trans_format("log10", function(x) as.character(10 ^ x))) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

## 2.5 Landwirtschaft (2/4) {.smaller}

```{r agriculture_plot2_faceted_scatter_normd}
# Normalize variable CO2 emissions to same scale as agricultural land [%]
data_agriculture_normd <- data_agriculture %>%
  group_by(`Country Name`) %>%
  mutate(
    co2_min = min(EN.GHG.CO2.MT.CE.AR5),
    co2_max = max(EN.GHG.CO2.MT.CE.AR5),
    co2_normd = ((EN.GHG.CO2.MT.CE.AR5 - co2_min) / (co2_max - co2_min))
  ) %>%
  ungroup()

# Pivot normalized data_agriculture dataset to longer format for better visualization of two variables
data_agriculture_normd_long <- data_agriculture_normd %>%
  pivot_longer(cols = c(co2_normd, AG.LND.AGRI.ZS), names_to = "variable", values_to = "value") %>%
  mutate(value = case_when(
    variable == "co2_normd" ~ value * 100,  # used for [%] comparison
    variable == "AG.LND.AGRI.ZS" ~ value))

# Plotting with normalized data over time for each country
ggplot(data_agriculture_normd_long,
       aes(x = Year, y = value, group = interaction(`Country Name`, variable))) +
  geom_point(aes(fill = variable), alpha = 0.75, size = 2.5, shape = 21, stroke = 0.5) +  # or geom_line? (And is geom_line overdramatizing the (wrong?) outliers in CO2 emissions?)
  labs(title = "Zusammenhang zwischen CO2 Emissionen pro Kopf\nund landwirtschaftlicher Nutzfläche nach Ländern",
       x = "Jahr", y = "Anteil [%]", fill = "") +
  scale_fill_brewer(palette = "Set2",  # Colorblind-friendly palette
                     labels = c("Landwirtschaftliche Nutzfläche [%]",
                                "(Min, Max)-normalisierte CO2 Emissionen pro Kopf [%]")) +
  scale_x_continuous(breaks = c(2000, 2005, 2010, 2015, 2020),  # Text labels for x-axis
                     minor_breaks = seq(2000, 2021, by = 1)) +  # Add ticks
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x = element_text(hjust = 0.5)) +
  facet_wrap(~ factor(data_agriculture_normd_long$`Country Name`, levels = country_order_by_agrLand), ncol = 5)
```

## 2.5 Landwirtschaft (3/4) {.smaller}

Spielt hierbei die vollständige Landfläche des Landes eine Rolle?

``` {r surface_setup_and_agriculture_plot3_groupfaceted}
source("Scripts/functions.R")
data_grouped <- readRDS("Data/Processed/data_pregrouped.RDS")
data_grouped <- group_data(df = data_grouped,
                               grouping_by = "AG.SRF.TOTL.K2",
                               column_new = "surfaceArea",
                               quantile_labels = c("Very Small", "Small", "Moderate", "Large", "Very Large"))

data_surface <- left_join(data_agriculture, data_grouped, by = c("Country Name", "Country Code")) 
data_surface <- data_surface %>%
  select(`Country Name`, `Country Code`, Year, AG.LND.AGRI.ZS, AG.SRF.TOTL.K2, EN.GHG.CO2.MT.CE.AR5, cat_surfaceArea)

# Plot the initial comparison via scatterplot, this time in grouped categories. 
ggplot(data_surface, aes(x = log10(EN.GHG.CO2.MT.CE.AR5 + 1), y = AG.LND.AGRI.ZS)) +
  geom_point(aes(color = cat_surfaceArea), alpha = .7, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title = "Zusammenhang zwischen CO2 Emissionen pro Kopf\nund landwirtschaftlicher Nutzfläche",
       x = "Logarithmierte CO2 Emissionen pro Kopf [Mt]", y = "Landwirtschaftliche Nutzfläche [%]") +
  scale_color_brewer(palette = "Set2", name = "Landfläche") +  
  # TODO: SHOULDN'T THIS BE A SEQUENTIAL COLORGRADING -> WHICH PALETTE?
  # scale_x_log10(name = "CO2 Emissionen pro Kopf",
  #               breaks = log10(c(0, 10, 100, 1000, 10000, 100000) + 1),
  #               labels = c("0", "10", "100", "1.000", "10.000", "100.000")) +
  # scale_x_log10(breaks = trans_breaks("log10", function(x) 10 ^ x),
  #               labels = trans_format("log10", function(x) as.character(10 ^ x))) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.title = element_text(hjust = 0.5))

# Plot the initial comparison via scatterplot, faceted in grouped categories. 
ggplot(data_surface, aes(x = AG.LND.AGRI.ZS, y = log10(EN.GHG.CO2.MT.CE.AR5 + 1))) +
  geom_point(alpha = .3, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  labs(title = paste0("Zusammenhang zwischen CO2 Emissionen pro Kopf und landwirtschaftlicher Nutzfläche",
                      "\n", "für Ländergruppen bezüglich deren Landfläche"),
       x = "Landwirtschaftliche Nutzfläche [%]", y = "Logarithmierte CO2 Emissionen pro Kopf [Mt]") +
  scale_color_brewer(palette = "Set2", name = "Landfläche") +  
  # TODO: SHOULDN'T THIS BE A SEQUENTIAL COLORGRADING -> WHICH PALETTE?
  # scale_x_log10(name = "CO2 Emissionen pro Kopf",
  #               breaks = log10(c(0, 10, 100, 1000, 10000, 100000) + 1),
  #               labels = c("0", "10", "100", "1.000", "10.000", "100.000")) +
  # scale_x_log10(breaks = trans_breaks("log10", function(x) 10 ^ x),
  #               labels = trans_format("log10", function(x) as.character(10 ^ x))) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.title = element_text(hjust = 0.5)) +
  facet_wrap(~ data_surface$cat_surfaceArea)
```

## 2.5 Landwirtschaft (4/4) {.smaller}

``` {r agriculture_plot4_groupfaceted_normd}
# Join the long normalized data for visualization of two variables on y-axis
data_surface_normd_long <- (left_join(data_agriculture_normd_long,
                                      data_grouped,
                                      by = c("Country Name", "Country Code")))


# Plotting with normalized data over time for each group
plot_surface2_upper_row <- ggplot(data_surface_normd_long %>%
                                    filter(cat_surfaceArea %in% c("Very Small", "Small", "Moderate")),
                                  aes(x = Year,
                                      y = value,
                                      group = interaction(`Country Name`, variable))) +
  geom_point(alpha = 0.5, size = 2.5, shape = 21, stroke = 0.5, aes(fill = variable)) +
  geom_smooth(aes(group = variable), method = "lm", se = FALSE, color = "black", size = 1.5) +
  geom_smooth(aes(group = variable, color = variable), method = "lm", se = FALSE, size = 1, show.legend = FALSE) +
  labs(title = paste0("Zusammenhang zwischen CO2 Emissionen pro Kopf und landwirtschaftlicher Nutzfläche",
                      "\n", "für Ländergruppen bezüglich deren Landfläche"),
       y = "Anteil [%]") +
  scale_fill_brewer(palette = "Set2",  # Colorblind-friendly palette
                     labels = c("Landwirtschaftliche Nutzfläche [%]",
                                "(Min, Max)-normalisierte CO2 Emissionen pro Kopf [%]")) +
  scale_color_brewer(palette = "Set2") +
  scale_x_continuous(breaks = c(2000, 2005, 2010, 2015, 2020),  # Text labels for x-axis
                     minor_breaks = seq(2000, 2021, by = 1)) +  # Add ticks
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none",
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 12),
        axis.text.x = element_text(hjust = 0.5)) +
  facet_wrap(~ cat_surfaceArea, ncol = 3)

plot_surface2_bottom_row <- ggplot(data_surface_normd_long %>%
                                     filter(cat_surfaceArea %in% c("Large", "Very Large")),
                                   aes(x = Year,
                                       y = value,
                                       group = interaction(`Country Name`, variable))) +
  geom_point(alpha = 0.5, size = 2.5, shape = 21, stroke = 0.5, aes(fill = variable)) +
  geom_smooth(aes(group = variable), method = "lm", se = FALSE, color = "black", size = 1.5) +
  geom_smooth(aes(group = variable, color = variable), method = "lm", se = FALSE, size = 1, show.legend = FALSE) +
  labs(x = "Jahr", y = "Anteil [%]", color = "", fill = "") +
  scale_fill_brewer(palette = "Set2",  # Colorblind-friendly palette
                     labels = c("Landwirtschaftliche Nutzfläche [%]",
                                "(Min, Max)-normalisierte CO2 Emissionen pro Kopf [%]")) +
  scale_color_brewer(palette = "Set2", labels = c("A", "B")) +
  scale_x_continuous(breaks = c(2000, 2005, 2010, 2015, 2020),  # Text labels for x-axis
                     minor_breaks = seq(2000, 2021, by = 1)) +  # Add ticks
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x = element_text(hjust = 0.5)) +
  facet_wrap(~ cat_surfaceArea, ncol = 2)

plot_surface2_upper_row /
  (plot_spacer() + plot_surface2_bottom_row + plot_spacer() + plot_layout(widths = c(.5, 2, .5)))
```

## Zusammenfassung und Diskussion {.smaller}
