---
title: "Worldbank Tobacco Analysis"
author: "Leonie Mertes"
date: "`r Sys.Date()`"
output:
  html_document: default
---
## Introduction
This analysis explores the relationship of prevalence of current tobacco use (% of adults) prevalence to GDP per capita. Therefore we analyse the following leading question:
1. How does GDP per capita relate to the prevalence of current tobacco use (% of adults)?
<br>
<br>
```{r setup, message = FALSE, echo = FALSE, warning = FALSE}
# Load necessary libraries
library(here)
library(dplyr)
library(ggplot2)

# Load the dataset
data_merged <- readRDS(here("Data", "Processed", "data_merged.RDS"))
data_grouped <- readRDS(here("Data", "Processed", "data_grouped.RDS"))

data_processed <- left_join(data_merged, data_grouped, by = c("Country Name", "Country Code"))
data_tob_all <- data_processed %>%
  select(`Country Name`,
         `Country Code`,
         Year,
         NY.GDP.PCAP.PP.KD,
         SH.PRV.SMOK,
         cat_tob_usage)
```
<br>
### 1. Tobacco Prevalence and GDP
<br>
Check how many NA's in tobacco variable each Year (excluding years existing only of NA's)
<br>
```{r, message = FALSE, echo = FALSE, warning = FALSE, fig.align = "left"}
data_table <- data_tob_all %>%
  group_by(Year) %>%
  reframe(num_na = sum(is.na(SH.PRV.SMOK))) %>%
  filter(num_na < 25)
print(data_table)
```
<br>
Knowing there are only 25 countries to investigate, we only have data of the Years 2000, 2005, 2010, 1015, 1018, 2019 and 2020 on percentage of tabacco usage in the adult population of all 24 countries (excluding Aruba). We again adjust our dataframe.
<br>
```{r, echo = FALSE, message = FALSE, fig.align = "left"}
data_tob <- data_tob_all %>% 
  filter(Year %in% data_table$Year) %>% 
  filter(`Country Name` != "Aruba")
```

<br>
To take a look at the grouping of our data by tobacco usage, we will display a plot of the most recent data in Year 2020.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob %>%
  filter(Year == 2020) %>%
  ggplot(aes(`Country Name`, SH.PRV.SMOK, colour = cat_tob_usage)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    y = "Anteil der tabakkonsumierenden Erwachsenen",
    x = "Land",
    title = "Gruppierung der Daten im Jahr 2020",
    colour = "relativer Tabakkonsum"
  ) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 4
  )) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now we take a look at the Comparison of tobacco usage and GDP in 2020. Therefore we group by the relative tobacco usage.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob %>%
  filter(Year == 2020) %>%
  group_by(cat_tob_usage) %>%
  ggplot(aes(SH.PRV.SMOK, NY.GDP.PCAP.PP.KD, colour = cat_tob_usage)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil der tabakkonsumierenden Erwachsenen in %",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP im Jahr 2020",
    colour = "relativer Tabakkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now we view only one dot per category of relative tobacco usage.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob %>%
  filter(Year == 2020) %>%
  group_by(cat_tob_usage) %>%
  summarise(
    cat_tob_usage = cat_tob_usage,
    NY.GDP.PCAP.PP.KD = mean(NY.GDP.PCAP.PP.KD, na.rm = TRUE),
    SH.PRV.SMOK = mean(SH.PRV.SMOK, na.rm = TRUE)
  ) %>%
  ggplot(aes(SH.PRV.SMOK, NY.GDP.PCAP.PP.KD, colour = cat_tob_usage)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil der tabakkonsumierenden Erwachsenen in %",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP im Jahr 2020",
    colour = "relativer Tabakkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now take a look of the coherence of the variables without our extreme value of the group "Very Low".
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob %>%
  filter(Year == 2020) %>%
  group_by(cat_tob_usage) %>%
  summarise(
    cat_tob_usage = cat_tob_usage,
    NY.GDP.PCAP.PP.KD = mean(NY.GDP.PCAP.PP.KD, na.rm = TRUE),
    SH.PRV.SMOK = mean(SH.PRV.SMOK, na.rm = TRUE)
  ) %>%
  filter(cat_tob_usage != "Very Low") %>%
  ggplot(aes(SH.PRV.SMOK, NY.GDP.PCAP.PP.KD, colour = cat_tob_usage)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil der tabakkonsumierenden Erwachsenen in %",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP im Jahr 2020",
    colour = "relativer Tabakkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
To get a better overview we take a look at all years. First only on years we have data of their tobacco consumption on.
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob %>%
  group_by(Year, cat_tob_usage) %>%
  summarise(
    NY.GDP.PCAP.PP.KD = mean(NY.GDP.PCAP.PP.KD, na.rm = TRUE),
    SH.PRV.SMOK = mean(SH.PRV.SMOK, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(Year, NY.GDP.PCAP.PP.KD, colour = cat_tob_usage)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(aes(group = cat_tob_usage)) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 6
  )) +
  labs(
    x = "Jahr",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP",
    colour = "relativer Tabakkonsum"
  ) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Now over all years we have data on their countries BIP.
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob_all %>%
  group_by(Year, cat_tob_usage) %>%
  summarise(
    NY.GDP.PCAP.PP.KD = mean(NY.GDP.PCAP.PP.KD, na.rm = TRUE),
    SH.PRV.SMOK = mean(SH.PRV.SMOK, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(Year, NY.GDP.PCAP.PP.KD, colour = cat_tob_usage)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(aes(group = cat_tob_usage)) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    size = 6
  )) +
  labs(
    x = "Jahr",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP",
    colour = "relativer Tabakkonsum"
  ) +
  scale_colour_brewer(palette = "Set2")
```
<br>
But maybe better facett over year to show better the coherence of BIP and tobacco consumption.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob %>%
  group_by(cat_tob_usage, Year) %>%
  summarise(
    NY.GDP.PCAP.PP.KD = mean(NY.GDP.PCAP.PP.KD, na.rm = TRUE),
    SH.PRV.SMOK = mean(SH.PRV.SMOK, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(SH.PRV.SMOK, NY.GDP.PCAP.PP.KD, colour = cat_tob_usage)) +
  facet_wrap(~Year) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil der tabakkonsumierenden Erwachsenen in %",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP alle Jahre",
    colour = "relativer Tabakkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Nun aber nochmal ohne unser lineares Modell, da durch den Wert für unser sehr niedrigen Tabbakkonsum dieses nicht geeignet ist.
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob %>%
  group_by(cat_tob_usage, Year) %>%
  summarise(
    NY.GDP.PCAP.PP.KD = mean(NY.GDP.PCAP.PP.KD, na.rm = TRUE),
    SH.PRV.SMOK = mean(SH.PRV.SMOK, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(SH.PRV.SMOK, NY.GDP.PCAP.PP.KD, colour = cat_tob_usage)) +
  facet_wrap(~Year) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil der tabakkonsumierenden Erwachsenen in %",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP alle Jahre",
    colour = "relativer Tabakkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "loess", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Wenn wir nun für jede Gruppe alle Länder anzeigen lassen, sieht der plot so aus:
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob %>%
  group_by(Year) %>%
  ggplot(aes(SH.PRV.SMOK, NY.GDP.PCAP.PP.KD, colour = cat_tob_usage)) +
  facet_wrap(~Year) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil der tabakkonsumierenden Erwachsenen in %",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP alle Jahre",
    colour = "relativer Tabakkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "loess", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Hier fällt jedoch auf, dass der relative Tabakkonsum nicht immer gut eingeordnet ist, da Gruppen mit weniger Tabakkonsum tatsächlich jedoch mehr Tabak konsumieren.
Hier sollte man sich überlegen, ob es nicht besser ist die Gruppierung für jedes jahr neu zu bestimmen.
Dieses Ergebnis wollen wir dann in einem Barplot festhalten. Hier nun also erst mit der allgemeineren Gruppierung:
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob %>%
  group_by(cat_tob_usage) %>%
  summarise(
    NY.GDP.PCAP.PP.KD = mean(NY.GDP.PCAP.PP.KD, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = cat_tob_usage, y = NY.GDP.PCAP.PP.KD)) +
  geom_col(fill = "steelblue") +
  labs(
    x = "relativer Tabakkonsum",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP"
  )
```
<br>
Und nun mit neuer Gruppierung der Daten:
Grouping by tobacco values each year and saving this full dataset in data_tob_cat_year using function 2.2 group_data_year
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
data_tob_cat_year <- data_tob %>% 
  mutate(cat_tob_usage_per_year = rep(NA, nrow(data_tob)))

data_tob_cat_year <- for (year in as.character(data_table$Year)) {
  
  new_cat <- group_data_year(df = data_tob_cat_year, grouping_by = "SH.PRV.SMOK", year = year, column_new = "tob_usage", quantile_labels = c("Sehr gering", "Gering", "Mittel", "Viel", "Sehr Viel"))
  
  data_filtered <- data_tob_cat_year %>% 
    filter(Year == year) %>% 
    mutate(cat_tob_usage_per_year = new_cat[ , 3, drop = FALSE])
  
  data_tob_cat_year <- data_tob_cat_year %>% 
    full_join(data_filtered, by = c("Year", "Country Name", "Country Code", "NY.GDP.PCAP.PP.KD", "SH.PRV.SMOK", "cat_tob_usage")) %>% 
    mutate(cat_tob_usage_per_year = coalesce(cat_tob_usage_per_year.x, cat_tob_usage_per_year.y)) %>% 
    select(-cat_tob_usage_per_year.x, -cat_tob_usage_per_year.y)
}
```
<br>
Hierzu nun erst der nach Jahr gefacettete dotplot:
<br>
```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
# noch nicht bearbeitet, nur grundgerüst gecopypasted
data_tob %>%
  group_by(cat_tob_usage, Year) %>%
  summarise(
    NY.GDP.PCAP.PP.KD = mean(NY.GDP.PCAP.PP.KD, na.rm = TRUE),
    SH.PRV.SMOK = mean(SH.PRV.SMOK, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(SH.PRV.SMOK, NY.GDP.PCAP.PP.KD, colour = cat_tob_usage)) +
  facet_wrap(~Year) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    x = "Anteil der tabakkonsumierenden Erwachsenen in %",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP alle Jahre",
    colour = "relativer Tabakkonsum"
  ) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  scale_colour_brewer(palette = "Set2")
```
<br>
Und nun im Barplot:
<br>
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = "left"}
# noch nicht bearbeitet, nur grundgerüst gecopypasted
data_tob %>%
  group_by(cat_tob_usage) %>%
  summarise(
    NY.GDP.PCAP.PP.KD = mean(NY.GDP.PCAP.PP.KD, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = cat_tob_usage, y = NY.GDP.PCAP.PP.KD)) +
  geom_col(fill = "steelblue") +
  labs(
    x = "relativer Tabakkonsum",
    y = "BIP pro Kopf in $",
    title = "Tabakkonsum vs. BIP"
  )
```
<br>
und nun barplot faccetiert nach jahr, x: jedes jahr neu gruppiert nach tabak und dann bip geplottet
<br>
```{r}

```
<br>