---
title: "World Bank Tobacco Analysis"
author: "Robin Billinger"
date: "`r Sys.Date()`"
output:
  html_document:
    self-contained: true
    clean: true
    css: "../customstyle.scss"
---

## Introduction
This analysis explores relationships between indicators across countries such as GDP per capita and prevalence of current tobacco usage (% of adults) using World Bank data. The question to be investigated is:
<br><br>
1. How does **GDP per capita** relate to the **prevalence of current tobacco use (% of adults)**?
<br><br>
For this analysis, we consider the following indicators:

```{r tobacco_setup, echo = FALSE, message = FALSE}
# Load libraries
library(knitr)
library(here)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(tidyr)
library(patchwork)

# Globally center all plots while displaying knitted version in HTML
knitr::opts_chunk$set(fig.align = "center")

# Create dictionary table to get overview over variables to work with
# Variables:  GDP per capita, PPP (constant 2021 international $)
#             NY.GDP.PCAP.PP.KD
#
#             Prevalence of current tobacco use (% of adults)
#             SH.PRV.SMOK
dictionary <- data.frame(
  Variable = c("NY.GDP.PCAP.PP.KD",
               "SH.PRV.SMOK"),
  `Indicator Name` = c("GDP per capita, PPP (constant 2021 international $)",
                       "Prevalence of current tobacco use (% of adults)"),
  Definition = c(
    paste0("GDP per capita based on purchasing power parity (PPP). PPP GDP is gross ",
           "domestic product converted to international dollars using purchasing power ",
           "parity rates. An international dollar has the same purchasing power over ",
           "GDP as the U.S. dollar has in the United States. GDP at purchaser's prices ",
           "is the sum of gross value added by all resident producers in the country ",
           "plus any product taxes and minus any subsidies not included in the value of ",
           "the products. It is calculated without making deductions for depreciation ",
           "of fabricated assets or for depletion and degradation of natural resources. ",
           "Data are in constant 2021 international dollars."),
    paste0("The percentage of the population ages 15 years and over who currently use ",
           "any tobacco product (smoked and/or smokeless tobacco) on a daily or non-",
           "daily basis. Tobacco products include cigarettes, pipes, cigars, cigarillos,",
           " waterpipes (hookah, shisha), bidis, kretek, heated tobacco products, and ",
           "all forms of smokeless (oral and nasal) tobacco. Tobacco products exclude ",
           "e-cigarettes (which do not contain tobacco), “e-cigars”, “e-hookahs”, JUUL ",
           "and “e-pipes”. The rates are age-standardized to the WHO Standard ",
           "Population.")),
  check.names = FALSE
)
kable(dictionary, caption = "World Bank Indicators", align = c("c", "c", "c"))

# Load the 'data_merged' dataset
data_tobacco <- readRDS(here("Data", "Processed", "data_merged.RDS"))

# Filter relevant columns Country Name and Code, Year, NY.GDP.PCAP.PP.KD and SH.PRV.SMOK
data_tobacco <- data_tobacco %>%
  select(`Country Name`, `Country Code`, Year, NY.GDP.PCAP.PP.KD, SH.PRV.SMOK)
```

## Agenda
1.) GDP per capita and prevalence of current tobacco use
<br><br>
1.1.) Scatter plot of interested variables
<br><br>
1.2.) Scatter plot of interested variables; except outlier Qatar
<br><br>
1.3.) Scatter plot of interested variables with faceted years
<br><br>
1.4.) Scatter plot of interested variables with faceted years with density curves
<br><br>
1.5.) Kullback-Leibler divergence for tobacco usage
<br><br>
2.) Summary

### 1. GDP per capita and prevalence of current tobacco use
We analyze how the GDP per capita for the observed countries relates to the prevalence of current tobacco use, representing the percentage of adults currently consuming tobacco. To get an overview over the interested data and be able to evaluate future insights correctly, we start by looking at the data available to us.
<br><br>

```{r missing_check, echo=FALSE, message=FALSE, warning=FALSE}
# Check whether any missing values in interested data subset
paste0("Are there no missing values? - Answer: ", all(!is.na(data_tobacco)))
paste0("How many missing values are in the data regarding GDP per capita? - Answer: ",
       sum(is.na(data_tobacco$NY.GDP.PCAP.PP.KD)))
paste0("How many missing values are in the data regarding tobacco prevalence? - Answer: ",
       sum(is.na(data_tobacco$SH.PRV.SMOK)))

# Check whether any countries with completely missing data.
data_tobacco %>%
  group_by(`Country Name`) %>%
  count(SH.PRV.SMOK) %>%
  filter(is.na(SH.PRV.SMOK)) %>%
  filter(n == 22)
```
As we can see, there is no data available for Aruba during the 22 years of time stretching from 2000 to 2021. Therefore, we drop Aruba from our analysis. Further, we check if there are any years in which the data is missing for several countries at the same time.

```{r missing_filter, include=FALSE, message=FALSE, warning=FALSE}
# As data to "Aruba" only contains NAs, remove Aruba from dataset
data_tobacco <- data_tobacco[data_tobacco$`Country Name` != "Aruba", ]

# Check which years are given for each of the other countries
data_tobacco %>%
  group_by(`Country Name`) %>%
  group_by(!is.na(SH.PRV.SMOK)) %>%
  count(Year) %>%
  print(n = 22)
```

```{r missing_filter_ext, echo=FALSE, message=FALSE, warning=FALSE}
# As we only have data on each country for the years 2000, 2005, 2010, 2015, 2018, 2019,
# and 2020, we remove all missing years for each country
data_tobacco <-
  data_tobacco[data_tobacco$Year %in% c(2000, 2005, 2010, 2015, 2018, 2019, 2020), ]

data_tobacco %>%
  group_by(`Country Name`) %>%
  group_by(!is.na(SH.PRV.SMOK)) %>%
  count(Year) %>%
  print(n = 22)
```
We recognize that the data available is actually only the years 2000, 2005, 2010, 2015, 2018, 2019, and 2020 for the remaining 24 countries. Therefore, that will be the years we are considering when moving forward.

#### 1.1. Scatter plot of interested variables

```{r scatter, echo=FALSE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
# Scatter plot of interested variables
ggplot(data_tobacco, aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, color = "Robust-Linear"),
              method = MASS::rlm,
              linewidth = 1, se = FALSE) +
  geom_smooth(aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, color = "Linear"),
              method = "lm", linewidth = 1, se = FALSE) +
  ggtitle("Zusammenhang zwischen Prävalenz des aktuellen Tabakkonsums und BIP pro Kopf") +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]",
                     breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(),
                     limits = c(0, 60)) +
  scale_y_continuous(name = "BIP pro Kopf [$]",
                     breaks = c(0, 25000, 50000, 75000, 100000, 125000),
                     minor_breaks = waiver(),
                     limits = c(0, 125000)) +
  scale_color_brewer(name = "", palette = "Set2") +
  theme_bw() +
  theme(axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```

We observe a slight trend within the relationship while the direction of this relationship seems to be heavily influenced by the high GDP per capita and low tobacco prevalence outliers in the upper left hand area of the plot according to the different regression models used.
<br>
As we are looking at the data without any consideration of time, let's check whether those data points all belong to the same country and, if that is actually the case, what the relationship would look like without this one-country-bias.

#### 1.2. Scatter plot of interested variables; except outlier Qatar

```{r scatter_withoutOutlier, echo=FALSE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
# Investigate the y-axis outlier - are they all from the same country?
data_tobacco %>% filter(NY.GDP.PCAP.PP.KD > 80000)
# Indeed, all the data points originate out of Qatar

# Scatter plot of interested variables; except outlier Qatar
ggplot(data_tobacco[data_tobacco$`Country Code` != "QAT", ],
       aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, color = "Robust-Linear"),
              method = MASS::rlm, linewidth = 1, se = FALSE) +
  geom_smooth(aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, color = "Linear"),
              method = "lm", linewidth = 1, se = FALSE) +
  ggtitle("Zusammenhang zwischen Prävalenz des aktuellen Tabakkonsums und BIP pro Kopf") +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]",
                     breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(),
                     limits = c(0, 60)) +
  scale_y_continuous(name = "BIP pro Kopf [$]",
                     breaks = c(0, 25000, 50000, 75000, 100000, 125000),
                     minor_breaks = waiver(),
                     limits = c(0, 125000)) +
  scale_color_brewer(name = "", palette = "Set2") +
  theme_bw() +
  theme(axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```

The removal of the Qatari data points leads us to slightly positive relationships for both, the linear as well as the robust-linear regression, visualized by the basically parallel straights divided only by marginal vertical differences.
<br>
Keeping this influence of the outliers in mind, we want to do the small intermediate step to check whether the linear relationship is actually represented in each of the observable years and introduce another form of regression to get deeper understanding of how much the linear representation is appropriate.

#### 1.3. Scatter plot of interested variables with faceted years

```{r faceted_scatter, fig.height=6, fig.width=10, echo=FALSE, message=FALSE, warning=FALSE}
# Scatter plot of interested variables with faceted years
ggplot(data_tobacco, aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD)) +
  geom_point(alpha = 0.5) +
  geom_smooth(data = data_tobacco, aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD,
                                       group = Year, color = "Robust-Linear"),
              method = MASS::rlm, linewidth = 1, se = FALSE) +
  geom_smooth(data = data_tobacco,
              aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, group = Year, color = "Linear"),
              method = "lm", linewidth = 1, se = FALSE) +
  geom_smooth(data = data_tobacco, aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD,
                                       group = Year, color = "Lokal-Polynomiell"),
              method = "loess", linewidth = 1, se = FALSE) +
  ggtitle("Zusammenhang zwischen Prävalenz des aktuellen Tabakkonsums und BIP pro Kopf") +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]",
                     breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(),
                     limits = c(0, 60)) +
  scale_y_continuous(name = "BIP pro Kopf [$]",
                     breaks = c(0, 25000, 50000, 75000, 100000, 125000),
                     minor_breaks = waiver(),
                     limits = c(0, 125000)) +
  scale_color_manual(name = "",
                     values = c("Linear" = brewer.pal(3, "Set2")[[1]],
                                "Robust-Linear" = brewer.pal(3, "Set2")[[2]],
                                "Lokal-Polynomiell" = brewer.pal(3, "Set2")[[3]])) +
  theme_bw() +
  theme(axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ Year, ncol = 4)
```

It seems that the Qatari data point in each of the years is still highly influential on the linear relationship. However, the relationship develops to more synchronized behaviour between the robust and non-robust relationship, with the non-robust changing from slightly negative to slightly positive.
<br>
But looking at the distribution of the data points within each facet, something very interesting is happening. With the higher density of points in the horizontal middle of the cloud and the simultaneous deviation of some of those points in the higher GDP direction, we can recognize a presumably new relationship we want to visualize next.

#### 1.4. Scatter plot of interested variables with faceted years with density curves

```{r faceted_scatter_density, fig.height=8, fig.width=12, echo=FALSE, message=FALSE, warning=FALSE}
# Prepare the density visualization:
# Calculate mean and standard deviation for each year
data_tobacco_param <- data_tobacco %>%
  group_by(Year) %>%
  summarize(SH.PRV.SMOK_mean = mean(SH.PRV.SMOK),
            SH.PRV.SMOK_sd = sd(SH.PRV.SMOK)) %>%
  ungroup()

# Generate evenly-spaced values of x and compute density for each year
data_tobacco_normdensity <- data_tobacco_param %>%
  rowwise() %>%
  mutate(
    # Create 100 evenly-spaced points in-between the lowest and highest possible x value
    norm_x_val = list(seq(0, 100, length.out = 101)),
    # Compute the corresponding PDF values
    norm_y_val = list(dnorm(norm_x_val,
                            mean = SH.PRV.SMOK_mean,
                            sd = SH.PRV.SMOK_sd))) %>%
  # Convert the lists into columns
  unnest(c(norm_x_val, norm_y_val))

# Remove rows with values of calculated normal distribution that are larger than 60%
# (i.e. 60% > max(SH.PRV.SMOK))
data_tobacco_normdensity <- data_tobacco_normdensity %>%
  filter(norm_x_val <= 60)

# Scatter plot of interested variables faceted years with density curves
# Plot the upper row of the data points for the assembled tobacco plotting
plot_tobacco1_data_upper <-
  ggplot(data_tobacco %>% filter(Year %in% c(2000, 2005, 2010, 2015))) +
  geom_point(aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, group = Year), alpha = 0.5) +
  geom_smooth(data = data_tobacco %>% filter(Year %in% c(2000, 2005, 2010, 2015)),
              aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, group = Year),
              method = MASS::rlm, linewidth = 1, se = FALSE, color = "black") +
  ggtitle("Prävalenz des Tabakkonsums vs. BIP pro Kopf") +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]",
                     breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(),
                     limits = c(0, 60)) +
  scale_y_continuous(name = "BIP pro Kopf [$]",
                     breaks = c(0, 25000, 50000, 75000, 100000, 125000),
                     minor_breaks = waiver(),
                     limits = c(0, 125000)) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 14),
        plot.margin = margin(0, 0, 0, 0),
        plot.title = element_text(hjust = 0.5, size = 18),
        strip.text = element_text(size = 12)) +
  facet_wrap(~ Year, ncol = 4)

# Plot the bottom row of the data points for the assembled tobacco plotting
plot_tobacco1_data_bottom <-
  ggplot(data_tobacco %>% filter(Year %in% c(2018, 2019, 2020))) +
  geom_point(aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, group = Year), alpha = 0.5) +
  geom_smooth(data = data_tobacco %>% filter(Year %in% c(2018, 2019, 2020)),
              aes(x = SH.PRV.SMOK, y = NY.GDP.PCAP.PP.KD, group = Year),
              method = MASS::rlm, linewidth = 1, se = FALSE, color = "black") +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]",
                     breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(),
                     limits = c(0, 60)) +
  scale_y_continuous(name = "BIP pro Kopf [$]",
                     breaks = c(0, 25000, 50000, 75000, 100000, 125000),
                     minor_breaks = waiver(),
                     limits = c(0, 125000)) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 14),
        legend.position = "bottom", 
        plot.margin = margin(0, 0, 0, 0),
        strip.text = element_text(size = 12)) +
  facet_wrap(~ Year)

# Add overlayed normal distribution curves to the assembled plot in the data displays
# Plot the upper row of the density curves for the assembled tobacco plotting
plot_tobacco1_density_upper <-
  ggplot() +
  geom_line(data = data_tobacco_normdensity %>%
              filter(Year %in% c(2000, 2005, 2010, 2015)),
            aes(x = norm_x_val, y = norm_y_val, group = Year),
            color = brewer.pal(4, "Set2")[[3]], linewidth = 1, inherit.aes = FALSE) +
  geom_density(data = data_tobacco %>%
                 filter(Year %in% c(2000, 2005, 2010, 2015)),
               aes(x = SH.PRV.SMOK, y = after_stat(density), group = Year),
               color = brewer.pal(4, "Set2")[[2]], linewidth = 1, bw = 3) +
  geom_vline(data = data_tobacco_normdensity %>%
               filter(Year %in% c(2000, 2005, 2010, 2015)),
             aes(xintercept = SH.PRV.SMOK_mean, group = Year),
             color = "black", linetype = "dashed") +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]",
                     breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(),
                     limits = c(0, 60)) +
  scale_y_continuous(name = "Dichte",
                     breaks = c(0, 0.025, 0.05),
                     minor_breaks = waiver(),
                     limits = c(0, 0.06)) +
  theme_bw() +
  theme(axis.text = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.title.y = element_text(margin = margin(r = 10), size = 14),
        plot.margin = margin(0, 0, 0, 0),
        strip.text = element_blank()) +
  facet_wrap(~ Year, ncol = 4)

# Plot the bottom row of the density curves for the assembled tobacco plotting
plot_tobacco1_density_bottom <-
  ggplot() +
  geom_line(data = data_tobacco_normdensity %>% filter(Year %in% c(2018, 2019, 2020)),
            aes(x = norm_x_val, y = norm_y_val, group = Year, color = "Normalverteilung"),
            linewidth = 1, inherit.aes = FALSE) +
  geom_density(data = data_tobacco %>% filter(Year %in% c(2018, 2019, 2020)),
               aes(x = SH.PRV.SMOK, y = after_stat(density), group = Year),
               color = brewer.pal(3, "Set2")[[2]], linewidth = 1, bw = 3,
               show.legend = FALSE) +
  # Add an invisible line by framing dummy line to show 'Kerndichteschaetzer' with line
  # shape instead of hollow square in legend
  geom_line(data = data.frame(x = c(0,0), y = c(0,0)),
            aes(x = x, y = y, color = "Kerndichteschätzer (Bandbreite = 3)"),
            linewidth = 1) +
  # Add an invisible line by framing dummy line to show 'Robust-Linear' with line
  # shape in the single legend below the whole assembled plot
  geom_line(data = data.frame(x = c(0,0), y = c(0,0)),
            aes(x = x, y = y, color = "Robust-Linearer Zusammenhang"),
            linewidth = 1) +
  geom_vline(data = data_tobacco_normdensity %>% filter(Year %in% c(2018, 2019, 2020)),
             aes(xintercept = SH.PRV.SMOK_mean, group = Year),
             color = "black", linetype = "dashed") +
  scale_x_continuous(name = "Prävalenz des aktuellen Tabakkonsums [%]",
                     breaks = c(0, 10, 20, 30, 40, 50, 60),
                     minor_breaks = waiver(),
                     limits = c(0, 60)) +
  scale_y_continuous(name = "Dichte",
                     breaks = c(0, 0.025, 0.05),
                     minor_breaks = waiver(),
                     limits = c(0, 0.06)) +
  scale_color_manual(
    name = "",
    values = c("Kerndichteschätzer (Bandbreite = 3)" = brewer.pal(3, "Set2")[[2]],
               "Normalverteilung" = brewer.pal(3, "Set2")[[3]],
               "Robust-Linearer Zusammenhang" = "black")) +
  theme_bw() +
  theme(axis.text = element_text(size = 12),
        axis.title.x = element_text(hjust = 0.85, margin = margin(t = 10), size = 14),
        axis.title.y = element_text(margin = margin(r = 10), size = 14),
        legend.position = c(0.667, -1.75),
        legend.direction = "horizontal",
        legend.text = element_text(size = 14),
        plot.margin = margin(0, 0, 0, 0),
        strip.text = element_blank()) +
  facet_wrap(~ Year)


# Assemble the plot
presentation_tobacco_plot2 <-
  (plot_tobacco1_data_upper /
    plot_tobacco1_density_upper /
    plot_spacer() /
    (plot_tobacco1_data_bottom + plot_spacer() + plot_layout(widths = c(3, 1))) /
    (plot_tobacco1_density_bottom + plot_spacer() + plot_layout(widths = c(3, 1))) +
    plot_layout(heights = c(1.9, 0.4, 0.1, 1.9, 0.4, 0.3), widths = c(1, 1, 1, 1))) &
  theme(plot.margin = margin(0, 0, 0, 0))

presentation_tobacco_plot2
```

We recognize the basically neutral relation in the year 2000 changing over time. Each year the trend gains slight positive increase, meaning over time, it changes towards the relationship of the more the country's GDP per capita, the higher its prevalence of tobacco use among adults. Further, it gets obvious that the general distribution of the data points moves to the left on the x-axis, so generally smaller occurrences of tobacco prevalence as time goes by. The same phenomenon can be observed when looking at the dashed vertical line in the density visualization below the scatter plots, as the mean decreases from each observed year to the next. Having said that, we recognize slightly contrary movement on y-axis, meaning higher GDP per capita for some of the countries, especially the points in the middle of the facets increase in GDP per capita. 
<br>
Moreover, the kernel density estimation seems to assimilate towards the normal distribution around the observed datas' means and standard deviations comparing the start and end of the timeframe.

#### 1.5. Kullback-Leibler divergence for tobacco usage

This can be proven by calculating the Kullback-Leibler divergence for each of the years, comparing the KDE with the underlying normal distribution to check the accuracy of describing our empirical distribution by normal distribution.

```{r kl_divergence, echo=FALSE, fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
# Source the functions.R script to access the 'kl_div_from_dnorm' function using 'here'
source(here("Scripts", "functions.R"))

# Compute Kullback-Leibler divergences for each year with differing kernel bandwidths
data_tobacco_kld <-
  kl_div_from_dnorm(data_tobacco, "SH.PRV.SMOK", kernel_bw = 2) %>%
  left_join(kl_div_from_dnorm(data_tobacco, "SH.PRV.SMOK", kernel_bw = 3)) %>%
  left_join(kl_div_from_dnorm(data_tobacco, "SH.PRV.SMOK", kernel_bw = 4)) %>%
  left_join(kl_div_from_dnorm(data_tobacco, "SH.PRV.SMOK", kernel_bw = 5))

colnames(data_tobacco_kld) <- gsub(pattern = "BW = ",
                                   replacement = "Breite = ",
                                   colnames(data_tobacco_kld),
                                   ignore.case = FALSE)
data_tobacco_kld <- data_tobacco_kld %>% rename("Jahr" = Year)

# Return table of Kullback-Leibler divergences for each year and eacg of the bandwidths
kable(data_tobacco_kld)
```

As we can see, there are clear differences between 2000 and 2020, regardless of granularity determined by the chosen bandwidth. Compared to the start of the time span, the empirical  distribution clearly attunes closer to the normal distribution at the end. However, the closest for each of the bandwidths is always 2010, although being very close for the small bandwidths (2, 3). For the bandwidth of two it is also the only one which decreases from 2010 onwards till the end. The wider bandwidths on the other hand, which smooth out the individual data points more with the risk of over-smoothing, decrease only for the first three facets and then increase again, all be it with marginal differences from 2018 to 2020. <br><br>

### 2. Summary

In summary though, seeing the small values of KL divergence in general, the marginal differences in the last years, and the shift of the mean towards the low-percentage area, we can conclude that the prevalence of tobacco usage among adults nears a normally distributed data cloud. After a clear convergence in the early 2000s, the convergence stagnated slightly over the last years tending to a backwards development but not strong enough to make any assumptions for future developments. Still, with the general average consumption decreasing.
